---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Basics of pipes

```{r}
x <- 1L:10
mean(x)
```

We can pipe without pipes
```{r}
x ->.; mean(.)
```

Although we still have . as an object
```{r}
.
```

In nutshell, pipe can be seen as:

```{r}
rm(.)
local({x -> .; mean(.)})
```

Although there is a lot additional stuff, that allows pipe to 


```{r}
library(magrittr)
x %>% mean
```

There are many types of pipes

```{r}
library(pipeR)
x %>>% mean
```

# Other pipes from magrittr

%T>%: breaks piping sequence.

```{r}
iris %>% 
  plot %>% 
  summary
```

In the example below iris is forwarded to plot by %T>%. The pipe operator %T>% also informs the next function, summary, that it should take as an argument the object precedenting %T>%.

%$%: works like attach()

```{r}
iris %$% 
  cor.test(Sepal.Length, Petal.Length)
```

%$% does wonders when working with lists. Compare normal pipe:

```{r}
my_list <- list(a = 1L:4,
                b = 5L:7)

my_list %>% 
  getElement("a") %>% 
  mean
```

and %$% pipe

```{r}
my_list %$% 
  mean(a)
```

It can even replace some dplyr idioms as pull:

```{r}
iris %>% 
  dplyr::pull(Sepal.Length) %>% 
  mean
```

where %$% allows you to remove one pipe.

```{r}
iris %$% 
  mean(Sepal.Length)
```

There is also assignment operator %<>%, which I really do not like.


# Readable code is not always the most efficient code

Pipes can slow you:

```{r}
library(microbenchmark)

microbenchmark(mean(x),
               x %>% mean,
               x %>>% mean,
               times = 1e4)
```

Let's see this in a bigger picture

```{r}
# pipe_bench <- lapply(1L:4, function(i) {
#   x <- 1L:(10^i)
#   summary(microbenchmark(mean(x),
#                          x %>% mean,
#                          x %>>% mean,
#                          times = 1e6)) %>% 
#     data.frame() %>% 
#     mutate(vector_size = max(x))
# }) %>% bind_rows()
# save(pipe_bench, file = "pipe_bench.RData")

load("pipe_bench.RData")
library(ggplot2)
ggplot(pipe_bench, aes(x = vector_size, y = mean, color = expr)) +
  geom_point(size = 4) +
  geom_line() +
  scale_x_log10() +
  theme_bw()
```

# Pipes and operators

Pipes are versatile

```{r}
x <- 1L:10
x %>% mean
x %>% mean()
x %>% mean(.)
```

The price for versatility is ambiguity, especially with operators

```{r}
vec <- 1L:10
y <- vec - 1
mean(y)
```

The intuitive solution fails:
```{r}
vec - 1 %>% mean
```

The following works properly and are synomous...
```{r}
(vec - 1) %>% mean
{vec - 1} %>% mean
```

...but be careful, because '{' and '(' are not equal.

```{r}
vec %>% {. - 1} %>% mean # yes
vec %>% (. - 1) %>% mean # no
```

The following structures are always wrong:
```{r}
vec %>% (.) - 1 %>% mean
vec %>% {.} - 1 %>% mean
```

And following examples are also wrong, but for a different reason

```{r}
vec %>% mean(. - 1)
vec  %>%  mean((. - 1)) 
vec  %>%  mean({. - 1})
```

Here, pipe forwards vec to mean() in a peculariar way. vec is a first argument, and {. - 1} is treated as an argument to trim.

We will study this behavior in depth using a simpler example:

```{r}
simple_fun <- function(x, y = 1, z = 1) {
  list(x, y, z)
}

(vec - 1) %>% simple_fun
```
```{r}
vec %>% simple_fun({. - 1})
```

If we use '{' or '(' inside the function called after pipe, it will automatically switch to the second argument. 

Of course, the most magrittr way to use operators is to rely on pipe-friendly aliases as 'add':

```{r}
vec %>% 
  add(-1) %>% 
  mean
```

# Customizing my shortcuts

```
remotes::install_github("michbur/addinexamplesWV")
```

Then go for Tools/Addins, choose you favorite pipes, add shortcuts (I like ctrl+shift+,) and enjoy piping!
